---
title: "Hands-on Exercise 1: ggplot using R"
categories: hands-on exercise
description: |
  Creating 1st data visualisation using ggplot package in R.
author:
  - name: Liang Minghao
    affiliation: SMU, MITB
    affiliation_url: https://scis.smu.edu.sg/master-it-business
date: "13 Jan, 2023"
preview:
output:
  distill::distill_article:
    self_contained: false
    toc: true
    toc_depth: 3
---

```{r global option, include = FALSE}
knitr::opts_chunk$set(message = FALSE, fig.retina = 3)
```
---

## Getting Started

### Import libraries & read sample data

```{r}
library(ggplot2)
library(tidyverse)
```

### Load Data

```{r}
d <- read.csv('/Users/minghaooo/Documents/Term 4/ISSS608-Visual Analytics/inclass-1/Exam_data.csv')
```

### Plot the average maths results of each class

```{r}
options(repr.plot.width = 5, repr.plot.height =2)
ggplot(d, aes(x=CLASS, y=MATHS),height = 10 , width = 50) + geom_bar(stat = "summary",fun = "mean")
```

### Working with facet

```{r}
# Facet by rows
ggplot(d, aes(x=CLASS, y=MATHS)) + geom_boxplot()+facet_grid(rows=vars(GENDER))

#Facet by columns
ggplot(d, aes(x=CLASS, y=MATHS)) + geom_boxplot()+facet_grid(cols=vars(GENDER))

#Facet Grid: rows + columns
ggplot(d, aes(x=GENDER, y=MATHS)) + geom_boxplot()+facet_grid(rows=vars(GENDER), cols=vars(CLASS))
```

### Working with theme

```{r}
ggplot(data=d, 
       aes(x=RACE)) +
  geom_bar() +
  coord_flip() +
  theme_minimal()+
  theme(panel.background = element_rect(fill = "lightblue",colour = "white"),
        panel.grid.major = element_line(size = 0.5, linetype = 'solid',colour = "white"), 
        panel.grid.minor = element_line(size = 0.25, linetype = 'solid',colour = "white")
        )
```

### The makeover design 1

-   Arrange bar charts in descending order

```{r}

d %>%
  mutate(RACE = fct_infreq(RACE)) %>%
  ggplot(aes(x=RACE)) + 
  geom_bar()+ 
  ylab('No. of Pupils')+
  theme(axis.title.y=element_text(angle=0))+
  geom_text(stat = 'count', 
            aes(label = paste(..count..,', ',round(..count../dim(d)[1]*100,digits=1), '%')),
            vjust = -0.5,
            size=3
            )
```

### The makeover design 2

-   Adding mean and median lines on the histogram plot
-   Change fill color and line color

```{r}
ggplot(data = d,
       aes(x=MATHS))+
  geom_histogram(bins = 20,
                 color = "black",
                 fill = "light blue") +
  geom_vline(xintercept = mean(d$MATHS),
             color = "red",
             linetype = "longdash") +
  annotate("text",x=mean(d$MATHS)-1,y=55, label="mean",color='red') +
  geom_vline(xintercept = median(d$MATHS),
             color = "black",
             linetype = "longdash")+
  annotate("text",x=median(d$MATHS)-1,y=50, label="median")
```

### The makeover design 3

-   The background histograms show the distribution of English scores for all pupils

```{r}
# duplicate a new dataset without gender
no_gender <- subset(d, select=-c(GENDER))

#fill diff. colour by gender
ggplot(d, aes(x=ENGLISH,fill=GENDER))+
  #plot background using duplicate dataset
  geom_histogram(data = no_gender, fill = "grey",binwidth = 2.5,alpha = 0.5)+
  #plot foreground
  geom_histogram(binwidth = 2.5, color = "black")+
  #facet by gender
  facet_wrap(~GENDER)
```

### The makeover design 4

-   A within group scatterplot with reference lines

```{r}
ggplot(data=d,
       aes(x = MATHS,y=ENGLISH)) +
  geom_point()+
  xlim(0,100)+
  ylim(0,100)+
  geom_vline(xintercept = 50, alpha=.5, linetype = "longdash") +
  geom_hline(yintercept = 50, alpha=.5, linetype = "longdash")
```
