---
title: "Hands-On Exercise 8"
categories: Hands-on exercise
description: |
  To model, analyse and visualise network data using R.
author:
  - name: Liang Minghao
    affiliation: SMU, MITB
    affiliation_url: https://scis.smu.edu.sg/master-it-business
date: "11 Mar, 2023"
preview: inclass_8_1.jpg
output:
  distill::distill_article:
    self_contained: false
    toc: true
    toc_depth: 3
---

# 1. Getting Started

## 1.1. Importing Libraries

In this hands-on exercise, four network data modelling and visualisation packages will be installed and launched: `igraph`, `tidygraph`, `ggraph` and `visNetwork.`

Beside these four packages, `tidyverse` and `lubridate`, an R package specially designed to handle and wrangling time data will be installed and launched too.

```{r}
pacman::p_load(igraph, tidygraph, ggraph, 
               visNetwork, lubridate, clock,
               tidyverse, graphlayouts)
```

## 1.2. Prepare Data

Edge data: GAStech-email_edges.csv which consists of two weeks of 9063 emails correspondances between 55 employees.

Nodes data: GAStech_email_nodes.csv which consist of the names, department and title of the 55 employees.

```{r}
GAStech_nodes <- read_csv("/Users/minghaooo/mtlmh34/ISSS608-VAA-mh/Data/s8/GAStech_email_node.csv")
GAStech_edges <- read_csv("/Users/minghaooo/mtlmh34/ISSS608-VAA-mh/Data/s8/GAStech_email_edge-v2.csv")
```

Show data.

```{r}
glimpse(GAStech_edges)
```

## 1.3. Wrangling time

-   Convert char type into date type;
-   Extract days in a week.

```{r}
GAStech_edges <- GAStech_edges %>%
  mutate(SendDate = dmy(SentDate)) %>%
  mutate(Weekday = wday(SentDate,
                        label = TRUE,
                        abbr = FALSE))
```

## 1.4. Wrangling Attributes

```{r}
GAStech_edges_aggregated <- GAStech_edges %>%
  filter(MainSubject == "Work related") %>%
  group_by(source, target, Weekday) %>%
    summarise(Weight = n()) %>%
  filter(source!=target) %>%
  filter(Weight > 1) %>%
  ungroup()
```

# 2. Create Networks

-   tbl_graph() creates a tbl_graph network object from nodes and edges data.
-   as_tbl_graph() converts network data and objects to a tbl_graph network.
    -   a node data.frame and an edge data.frame,
    -   data.frame, list, matrix from base,
    -   igraph from igraph,
    -   network from network,
    -   dendrogram and hclust from stats,
    -   Node from data.tree,
    -   phylo and evonet from ape,
    -   and graphNEL, graphAM, graphBAM from graph (in Bioconductor).

```{r}
GAStech_graph <- tbl_graph(nodes = GAStech_nodes,
                           edges = GAStech_edges_aggregated, 
                           directed = TRUE)
GAStech_graph
```

```{r}
# arrange edges by weight
GAStech_graph %>%
  activate(edges) %>%
  arrange(desc(Weight))
```

# 3. Visualization of networks

As in all network graph, there are three main aspects to a ggraph's network graph, they are: - Nodes, - Edges, - Layouts.

## 3.1. Default network graph

```{r}
ggraph(GAStech_graph) +
  geom_edge_link() +
  geom_node_point()
```

## 3.2. Changing theme

```{r}
g <- ggraph(GAStech_graph) + 
  geom_edge_link(aes()) +
  geom_node_point(aes())

g + theme_graph()
```

## 3.3. Changing colors

```{r}
g <- ggraph(GAStech_graph) + 
  geom_edge_link(aes(colour = 'grey50')) +
  geom_node_point(aes(colour = 'grey40'))

g + theme_graph(background = 'grey10',
                text_colour = 'white')
```

## 3.4. Changing layouts

Layouts for `ggraph()`: star, circle, nicely (default), dh, gem, graphopt, grid, mds, spahere, randomly, fr, kk, drl and lgl.

```{r}
g <- ggraph(GAStech_graph, 
            layout = "kk") +
  geom_edge_link(aes()) +
  geom_node_point(aes())

g + theme_graph()
```

## 3.5. Modifying nodes

```{r}
g <- ggraph(GAStech_graph, 
            layout = "nicely") + 
  geom_edge_link(aes()) +
  geom_node_point(aes(colour = Department, 
                      # colour each node by referring to their respective departments
                      size = 3))

g + theme_graph()
```

## 3.6. Modifying edges

```{r}
g <- ggraph(GAStech_graph, 
            layout = "nicely") +
  geom_edge_link(aes(width=Weight), 
                 # thickness of the edges will be mapped with the Weight variable
                 alpha=0.2) +
  scale_edge_width(range = c(0.1, 5)) +
  geom_node_point(aes(colour = Department), 
                  size = 3)

g + theme_graph()
```

## 3.7. Creating facet graphs

```{r}
set_graph_style()

g <- ggraph(GAStech_graph, 
            layout = "nicely") + 
  geom_edge_link(aes(width=Weight), 
                 alpha=0.2) +
  scale_edge_width(range = c(0.1, 5)) +
  geom_node_point(aes(colour = Department), 
                  size = 2) +
  theme(legend.position = 'bottom')
  
g + facet_edges(~Weekday)
```

## 3.8. Creating a framed facet graph

```{r}
set_graph_style() 

g <- ggraph(GAStech_graph, 
            layout = "nicely") + 
  geom_edge_link(aes(width=Weight), 
                 alpha=0.2) +
  scale_edge_width(range = c(0.1, 5)) +
  geom_node_point(aes(colour = Department), 
                  size = 2)
  
g + facet_edges(~Weekday) +
  th_foreground(foreground = "grey80",  
                border = TRUE) +
  theme(legend.position = 'bottom')
```
