---
title: "In-Class Exercise 4"
categories: in-class exercise
description: |
  R: Fundamentals of Visual Analytics
author:
  - name: Liang Minghao
    affiliation: SMU, MITB
    affiliation_url: https://scis.smu.edu.sg/master-it-business
date: "Feb 04 2023"
date modified: "`r Sys.Date()`"
format: html
execute: 
  echo: true
  eval: true
  warning: false
preview: 
output:
  distill::distill_article:
    self_contained: false
    toc: true
    toc_depth: 3
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Load Packages and Exam Data

```{r}
pacman::p_load(plotly, DT, patchwork, ggstatsplot, tidyverse)
exam_data <- read_csv('../Data/Exam_data.csv')
```

## 1. Viz Using `Plotly`

### Using `Plotly` API Directly

```{r}
plot_ly(data = exam_data, 
        x = ~ENGLISH, 
        y = ~MATHS, 
        color = ~RACE)
```

### Use `ggplotly()` Wrapper

create normal `ggplot()` first, and use `ggplotly()` to wrap over the plot object.

```{r}
p <- ggplot(data=exam_data, 
            aes(x = MATHS,
                y = ENGLISH)) +
  geom_point(size = 1) +
  coord_cartesian(xlim=c(0,100),
                  ylim=c(0,100))
ggplotly(p) #<<
```

## 2. Visual Statistical Analysis with `ggstatsplot`

### Two-sample mean test: `ggbetweenstats()`

```{r}
ggbetweenstats(
  data = exam_data,
  x = GENDER, 
  y = MATHS,
  type = "p",
  messages = FALSE
)
```

### Significant Test of Correlation: `ggscatterstats()`

```{r}
ggscatterstats(
  data = exam_data,
  x = MATHS,
  y = ENGLISH,
  marginal = TRUE, #marginal histogram
  )
```

## 3. Visualising Models

```{r}
pacman::p_load(readxl, performance, parameters, see, gtsummary)
```

```{r}
car_resale <- read_xls('../Data/ToyotaCorolla.xls')
```

Linear regression model

```{r}
model <- lm(Price ~ Age_08_04 + Mfg_Year + KM + 
              Weight + Guarantee_Period, data = car_resale)

#use gtsummary to show tabulated result of the variables
tbl_regression(model)
```

#### Visualizing multi-collinearity

```{r}
check_c <- check_collinearity(model)
plot(check_c)
```

New model, with the highly correlated variable taken out:

```{r}
model1 <- lm(Price ~ Age_08_04 + KM + 
              Weight + Guarantee_Period, data = car_resale)
check_n <- check_normality(model1)
plot(check_n)
```

#### Complete Dignostic of Model

```{r,fig.width=12,fig.height=9}
check_model(model1)
```

#### Visualising Regression Parameters

```{r}
ggcoefstats(model1, 
            output = "plot")
```

## 4. Visualising Uncertainty

Calculate no. of rows, mean and s.d., aggregated by RACE.

```{r}
my_sum <- exam_data %>%
  group_by(RACE) %>%
  summarise(
    n=n(),
    mean=mean(MATHS),
    sd=sd(MATHS)
    ) %>%
  mutate(se=sd/sqrt(n-1))
my_sum
```

```{r}
ggplot(my_sum) +
  geom_errorbar(
    aes(x=RACE, 
        ymin=mean-se, 
        ymax=mean+se), 
    width=0.2, 
    colour="black", 
    alpha=0.9, 
    size=0.5) +
  geom_point(aes
           (x=RACE, 
            y=mean), 
           stat="identity", 
           color="red",
           size = 1.5,
           alpha= 0.7) +
  ggtitle("Standard error of mean 
          maths score by race")
```
